---
title: "Data Visualization"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  eval = TRUE,
  message = FALSE,
  warning = FALSE)
```

## Introduction to data visualisation 
### Load the packages 
```{r}
library(tidyverse)
library(learnr)
library(ggplot2)
library(gridExtra)
```

An interesting chocolate data set with diverse ingredients in chocolates. Firstly, library `readr` and then use `read_csv()` function to import the chocolate data set.
```{r}
library(readr)
chocolates <- read_csv("data/chocolates.csv")
```

```{r echo = FALSE}
chocolates <- chocolates %>%
  mutate(Contain_Chol = ifelse(Chol_mg > 0, TRUE, FALSE)) %>% 
  mutate(Type = factor(Type))
```

* Different chocolates in rows are our observations and ingredients are variables.
* Our variables include fiber, fat, calories countries and etc...
```{r}
head(chocolates) 
```

* Use `glimpse()` function to have a look at our data type relative to different variables; The dimension of the data can be seen at the top left 
* You can also use `dim()` function to know the dimension of the data: 88 observations and 15 variables 
* Use `names()` to see the relevant variables  
* Data type: character, factor, double and logical 
```{r}
glimpse(chocolates) 
dim(chocolates)
names(chocolates)
```

### Make a plot by using chocolate data 
* Use `ggplot()` to visualize the data  
* Change the title and the name of x axis by using `ggtitle()` and `xlab()`
```{r}
ggplot(data = chocolates, aes(x = Sugars_g, y = Calories, color = Type)) + 
  geom_point() + 
  ggtitle("The relationship between calories and sugars relative to chocolate types") +
  xlab("Sugars")
```

## Scatter plot
* Using function `geom_point()` in the `ggplot2` package 
* Using mtcars as our data 

### What is mtcars?
* Motor Trend Car Road Tests -- the data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models).
* Use `help()` function to find more information 
* Important variables: 
    + mpg -- Miles/(US) gallon
    + cyl -- Number of cylinders
    + wt -- Weight (1000 lbs)
    + vs -- Engine (0 = V-shaped, 1 = straight)
    + am -- Transmission (0 = automatic, 1 = manual)
    + hp -- Gross horsepower
```{r}
glimpse(mtcars)
```

### Simple point plot 
The graph indicates that the cars with lower weight tend to use less petrol.
```{r}
ggplot(data = mtcars, aes(x = wt, y = mpg)) + 
  geom_point() 
```

### Color with discrete variable
* Use `factor()` to make `cyl` as a factor  
* `color` is in the `aes()` function 
* The function `labs()` can be used to easily update the legend title for a given aesthetic (fill, color, size, shape, line type, …)
* We can also change the text color and position of the legend as well

The graph demonstrates the miles of the cars travel relative to the number of cylinders the cars have. Heavy cars tend to have more cylinders and travel fewer distances than light cars with the same amount of gasoline. 
```{r}
p1 <- ggplot(data = mtcars, aes(x = wt, y = mpg, color = factor(cyl))) + 
  geom_point() +
  labs(color = "cyl")

p2 <- ggplot(data = mtcars, aes(x = wt, y = mpg, color = factor(cyl))) + 
  geom_point() +
  labs(color = "cyl") +
  theme(legend.position = "top",
        legend.title = element_text(color = "blue", size = 10, face = "bold"),
        legend.text = element_text(color = "blue", size = 10, face = "bold"))

grid.arrange(p1, p2, nrow = 1)
```

### Color with continuous variable
#### Question: What kind of car tends to have high horsepower?
`scale_color_gradient()` for sequential gradients between two colors
```{r}
ggplot(data = mtcars, aes(x = wt, y = mpg, color = hp)) + 
  geom_point() +
  labs(color = "hp") +
  scale_color_gradient(low="blue", high="red")
```

### Selecting point color 
```{r}
ggplot(data = mtcars, aes(x = wt, y = mpg)) + 
  geom_point(color = "blue")
```

### Selecting point shapes
```{r}
ggplot(data = mtcars, aes(x = wt, y = mpg, shape = factor(cyl))) + 
  geom_point() + 
  labs(shape = "cyl")
```

### Selecting point sizes
```{r}
ggplot(data = mtcars, aes(x = wt, y = mpg, size = factor(cyl))) + 
  geom_point() +
  labs(size = "cyl")
```

### Selecting point transparency
```{r}
ggplot(data = mtcars, aes(x = wt, y = mpg, alpha = factor(cyl))) + 
  geom_point() + 
  labs(alpha = "cyl")
```

### Combining color and shape 
#### What you can get from this graph?
```{r}
ggplot(data = mtcars, aes(x = wt, y = mpg, shape = factor(am), color = hp)) + 
  geom_point() +
  labs(x = "Weight (1000 lbs)", y = "Miles per gallon", color = "Gross horsepower") +
  scale_shape_discrete(name = "Transmission", labels = c("automatic", "manual"))
```

* Manual cars have high efficiency of using petrol 
* One manual car has the highest horsepower -- think about sports cars

### Adding a vertical line 
```{r}
ggplot(data = mtcars, aes(x = wt, y = mpg)) + 
  geom_point() +
  geom_vline(xintercept = 3)

ggplot(data = mtcars, aes(x = wt, y = mpg)) + 
  geom_point() +
  geom_vline(xintercept = 3, linetype = "dotted", 
             color = "blue", size = 1.5)
```

## Line graph 
### Simple line graph
* Using `mpg` as our data frame and use `help()` function to get more information 
* Fuel economy data from 1999 and 2008 for 38 popular models of car
* Using function `geom_line()`
```{r}
ggplot(data = mpg, aes(x = displ, y = hwy)) + 
  geom_line()
```

## Faceting 
### Using `facet_wrap()`
By using faceting, we can clearly identify the features of cars with different number of cylinders. 
```{r}
ggplot(data = mtcars, aes(x = wt, y = mpg)) + 
  geom_point() + 
  facet_wrap(~cyl)
```

### Using `facet_grid()`
Add facets categorized by multiple variables
```{r}
ggplot(data = mtcars) +
  geom_point(mapping = aes(x = wt, y = mpg)) + 
  facet_grid(cyl ~ am)
```

## Box plot, histogram, bar chart and scatter plot matrices
### Box plot 
* Change the background by using `theme_bw()` function 
* Add color corresponding to `drv` 
```{r}
ggplot(data = mpg, aes(x = drv, y = hwy, fill = drv)) + 
  geom_boxplot() +
  theme_bw()
```

### Histogram 
* Using `binwidth()` to adjust the width of each bin  
* we only need y axis in the `aes()` 
```{r}
ggplot(data = mpg, aes(hwy)) + 
  geom_histogram(binwidth = 0.6)
```

### Bar chart
```{r}
ggplot(data = mpg, aes(hwy)) + 
  geom_bar()
```

### Count of `drv` in bar chart 
```{r}
ggplot(data = mpg, aes(hwy, fill = drv)) + 
  geom_bar() +
  scale_fill_brewer(palette="Dark2")
```

### Compositional bar charts
When cars are moving using front wheels on highway, they tend to drive more miles per gallon instead of using all 4 wheels or only rear wheels. 
```{r}
ggplot(data = mpg, aes(hwy, fill = drv)) + 
  geom_bar(position="fill") +
  scale_fill_brewer(palette="Dark2")
```

### Side-by-side bar charts 
When cars are moving on the highway, they are more likely to use 4 wheels or just front wheels. 
```{r}
ggplot(data = mpg, aes(x = hwy, fill = drv)) + 
  geom_bar(position="dodge") +
  scale_fill_brewer(palette="Dark2")
```

### What can you get from this complicated graph? 
- Different types of cars use different wheels to move as they may have different functions 
- Small cars tend to use petrol more efficiently on highway than large cars 
```{r}
ggplot(data = mpg, aes(x = hwy, fill = drv)) + 
  geom_bar(position="dodge") +
  facet_wrap(~class) +
  scale_fill_brewer(palette="Dark2")
```

### Scatter plot Matrices 
* All variables are supposed to be numerical 
* Rotate the text of x axis in order to be seen clearly 
```{r}
library(GGally)
mtcars %>% 
  mutate(am = factor(am)) %>% 
  ggscatmat(data=.,color = "am", columns=c(1,3,4,5,6,7)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## Non-linear regression 
* Using span to control the "willingness" of the default loess smoother.
* The span is the fraction of points used to fit each local regression.
* Small numbers make a wigglier curve, larger numbers make a smoother curve.
```{r}
p3 <- ggplot(data = mpg, aes(x = displ, y = hwy)) + 
  geom_point() +
  geom_smooth()

p4 <- ggplot(data = mpg, aes(x = displ, y = hwy)) + 
  geom_point() +
  geom_smooth(span = 0.3)

grid.arrange(p3, p4, nrow = 1)
```

### Set `se = FALSE` to get rid of standard error
```{r}
ggplot(data = mpg, aes(x = displ, y = hwy)) + 
  geom_point() +
  geom_smooth(span = 0.3, se = FALSE)
```


